Categorical Predictor Embeddings in Actuarial Modeling

### 2021-03-09 manuscript/manuscript.Rmd
edits

The working example for our experiments is as follows: Given a set of claim characteristics, we predict the losses paid 
on the property coverage of the policy.

The data we use comes from the National Flood Insurance Program (NFIP) and is made available by the OpenFEMA initiative 
of the Federal Emergency Management Agency (FEMA)[^9]. Two datasets are made available by OpenFEMA: A policies dataset 
with exposure information, and a claims dataset with claims transactions, including paid amounts. Because there is no 
way to associate records of the two datasets, we are limited to fitting severity models on the claims dataset. While the 
complete dataset contains over two million transactions, for the purposes of our experiments we limit ourselves to data 
from 2000 to August 2019, which amounts to approximately 1.4 million claims. The dataset can be downloaded from 
Cellar[^10] The dataset includes a rich variety of variables, from occupancy type to coarse coordinates. For our models, 
we work with a few selected variables that represent continuous and discrete variables of low and high cardinalities, 
which we list in Table \@ref(tab:variables)

@@

```{r variables, results='asis', message=FALSE, echo = FALSE}
library(magrittr)
variables <- tibble::tribble(
  ~Variable, ~Type,
  "Building insurance coverage", "Numeric",
  "Basement enclosure type", "Categorical",
  "Number of floors in the insured building", "Categorical (binned in original dataset)",
  "Flood zone", "Categorical (high cardinality)",
  "Primary residence", "Categorical (binary indicator)"
)
knitr::kable(
  variables,
  caption = "Predicator variables.",
  booktabs = TRUE,
  table.env = "table",
  escape = FALSE,
  linesep = ""
) %>%
  kableExtra::kable_styling(latex_options = "hold_position")
```

@@

1. A GLM with gamma distribution and log link function
2. A linear model with categorical embeddings and an exponential output activation
3. A GLM with the categorical predictors replaced by the trained embeddings from Model 2.
4. A multilayer perceptron regression model

>>>>

The working example for our experiments is as follows: Given a set of claim characteristics, we predict the losses paid 
on the property coverage of the policy. The data we use comes from the National Flood Insurance Program (NFIP) and is 
made available by the OpenFEMA initiative 
of the Federal Emergency Management Agency (FEMA)[^9]. Two datasets are made available by OpenFEMA: A policies dataset 
with exposure information, and a claims dataset with claims transactions, including paid amounts. Because there is no 
way to associate records of the two datasets, we are limited to fitting severity models on the claims dataset. While the 
complete dataset contains over two million transactions, for the purposes of our experiments we limit ourselves to data 
from 2000 to August 2019, which amounts to approximately 1.4 million claims. The dataset can be downloaded from 
Cellar[^10] The dataset includes a rich variety of variables, from occupancy type to coarse coordinates. For our models, 
we work with a few selected variables that represent continuous and discrete variables of low and high cardinalities, 
which we list in Table \@ref(tab:variables)

@@

```{r variables, results='asis', message=FALSE, echo = FALSE}
library(magrittr)
variables <- tibble::tribble(
  ~Variable, ~Type,
  "Building insurance coverage", "Numeric",
  "Community Rating System Discount", "Numeric",
  "Basement enclosure type", "Categorical",
  "Occupancy Type", "Categorical",
  "Number of floors in the insured building", "Categorical (binned in original dataset)",
  "Flood zone", "Categorical (high cardinality)",
  "Primary residence", "Categorical (binary indicator)"
)
knitr::kable(
  variables,
  caption = "Predicator variables.",
  booktabs = TRUE,
  table.env = "table",
  escape = FALSE,
  linesep = ""
) %>%
  kableExtra::kable_styling(latex_options = "hold_position")
```

The response variable we take from the dataset is "Amount Paid on Building Claim", which is a numeric variable.

@@

1. A GLM with gamma distribution and log link function
2. A linear model with categorical embeddings and an exponential output activation
3. A GLM with the categorical predictors replaced by the trained embeddings from Model 2
4. A multilayer perceptron regression model
